---
title: "SkinPanel Rank"
author: "Francesco Gastaldello"
date: "12 July 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Skin Panel

The skin panel has been requested by Andrew Woodland and Elad Katz, members of the DGEM group at the Drug Discovery Unit. It's a collection of expresses genes in the skin from different databases: GTeX (<https://www.gtexportal.org/home/>) and Protein Atlas (<http://www.proteinatlas.org>). The numerical data express gene counts in the unit measure cited in the column names.

Lets load and subset the main file by considering only GTex and Protein Atlas columns:

```{r skin_panel_flatfile}
sp <- read.delim('Z:/SkinPanel/SkinPanel_kinases.csv')
sp <- sp[1:11]
```

Remove those columns that do not contain any information about gene expression.

```{r results='sp',error=F,warning=F}
apply(sp[,3:9],2,function(x){sum(x)==0})
# remove the columns with no data
sp <- subset(sp,select=-c(6:9))
knitr::kable(head(sp[,1:5]), format = "markdown")
```
## Is data normalized?

Check if data is normalized

```{r, fig.height=10, fig.width=16}
par(mfrow=c(1,3))
qqnorm(sp[,3],main = "GTex No Sun");qqline(sp[,3],col=2)
qqnorm(sp[,4],main = "GTex Sun");qqline(sp[,4],col=2)
qqnorm(sp[,5],main = "PA keratinocytes");qqline(sp[,5],col=2)
```
```{r echo=F}
par(mfrow=c(1,1))
```

Normalize and scale data:

```{r}
sp_norm <- sp
sp_norm[,3] <- log10(sp_norm[,3]/median(sp_norm[,3]))

# ignore the 0's from the column 4
sp_norm[,4] <- log10(sp_norm[,4]/median(sp_norm[,4][sp_norm[,4]>0]))
sp_norm[,5] <- log10(sp_norm[,5]/median(sp_norm[,5]))
```

Check data distribution:

```{r, fig.height=14, fig.width=16}
# check data distribution
par(mfrow=c(1,3))
hist(sp_norm[,3],main = "GTex No Sun")
hist(sp_norm[,4],main = "GTex Sun")
hist(sp_norm[,5],main = "PA keratinocytes")
```
```{r echo=F}
par(mfrow=c(1,1))
```

## Ranking
One othe the chemist in the teasm asked if it was possible to provide some sort of ranking for the genes.

Must remember to check for each column the max value and change any *-Inf* with *max_value*.

```{r}
sp_ranked <- sp_norm

changeInf <- function(d){
  max_value <- max(d)
  d <- replace(d,d == -Inf, max_value)
}
sp_ranked[,3] <- changeInf(sp_ranked[,3])
sp_ranked[,4] <- changeInf(sp_ranked[,4])
sp_ranked[,5] <- changeInf(sp_ranked[,5])
```

Rank data:

```{r}
sp_ranked[,8] <- rank(sp_ranked[,3],ties.method = "average")
colnames(sp_ranked)[8] <- "GTex No Sun Rank"
sp_ranked[,9] <-rank(sp_ranked[,4],ties.method = "average")
colnames(sp_ranked)[9] <- "GTex Sun Rank"
sp_ranked[,10] <-rank(sp_ranked[,5],ties.method = "average")
colnames(sp_ranked)[10] <- "PA keratinocytes Rank"
```


Calculate the **MEDIAN** of all ranks into a single variable:

```{r}
sp_ranked[,11] <- apply(sp_ranked[,8:10],1,median)
colnames(sp_ranked)[11] <-"MEDIAN Rank"
knitr::kable(head(sp_ranked[,c(1,2,8,9,10,11)]), format = "markdown")
```

